---
title: "Renters"
output: html_document
date: "2025-02-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(stargazer)
library(knitr)
library(kableExtra)
library(plm)
library(lme4)
library(sf)
library(leaflet)
library(spdep)
library(spatialreg)
library(sp)
library(terra)
library(tmap)
library(splm)
library(plm)
library(corrplot)
library(gridExtra)
library(car)
library(lmtest)
```

# Loading in Data

```{r}
df <- read.csv("C:/Users/timmy/Desktop/School Files/Capstone/Final CSVs/Final Spatial Data Natural Hazards.csv")
states_sf <- st_read("C:/Users/timmy/Desktop/School Files/Capstone/tl_2024_us_state.shp", promote_to_multi = FALSE)

df$State <- as.factor(df$State)
```

```{r}
#states_sf <- st_cast(states_sf, "POLYGON")

states_sf <- states_sf %>%
  rename(State = NAME)

# Temporarily save geometry and attributes separately
geometry <- st_geometry(states_sf)
attributes <- st_drop_geometry(states_sf)

# Expand attributes to include all years
expanded_attributes <- attributes %>%
  mutate(dummy = 1) %>%  
  expand_grid(Year = 2008:2019) %>%  
  select(-dummy)  

# Combine expanded attributes with original geometry
states_sf <- st_sf(expanded_attributes, geometry = rep(geometry, each = length(2008:2019)))
```

# Data Wrangling

```{r}
df_test <- df 
#df_test$Percent_white <- df_test$Percent_white * 100

#df_test$DAMAGE_PROPERTY <- df_test$DAMAGE_PROPERTY / 1000
#df_test$DAMAGE_CROPS <- df_test$DAMAGE_CROPS / 1000
```

```{r}
final_df <- states_sf %>%
  inner_join(df_test, by = c("State", "Year"))
```

```{r}
final_df <- final_df %>%
  filter(!State %in% c("Hawaii", "Alaska"))

final_df <- final_df %>%
  select(-total_damage_property.y, -total_damage_crops.y) %>%  
  rename(
    total_damage_property = total_damage_property.x,  
    total_damage_crops = total_damage_crops.x  
  )

#model_ready <- final_df %>%
  #arrange(Year, State)  %>% 
  #mutate(total_damage = total_damage_property + total_damage_crops) %>% 
  #mutate(td_per_cap = total_damage / Total_Residents)

model_ready <- final_df %>%
  arrange(Year, State)  %>% 
  mutate(total_damage = total_damage_property) %>% 
  mutate(td_per_cap = total_damage / Total_Residents)

model_ready <- model_ready %>%
  rename(
    homePrems = Home.Avg.Premium,
    rentersPrems = Renters.Avg.Premium,
    HealthcareExp = Health.Spending.per.Capita,
    Pop18_64 = Total_Pop_Dis,
    Uninsured = Percent.Adults.19.64.Uninsured,
    Medicaid = Percent,
    PriceDef = RegionalPriDef,
    Coastal = Is_Coastal,
    MedHomeIncome = Median.Income,
    AvgHomePrice = Avg_Home_Price,
    RiskIndex = National_Risk_Index_Score_Composite,
    ExpAnnualLoss = Expected.Annual.Loss...Score...Composite
  )
```

I redid my damage variable because, honestly, my previous method was pretty lazy. This time, I separated damage by type (property vs. crops) and by storm event. This ensures the model has access to the correct damage numbers. Additionally, based on this research paper https://www.nber.org/system/files/working_papers/w32579/w32579.pdf, I decided to split the damage into categories: those covered by typical homeowners insurance and those requiring separate coverage (earthquake, flood, etc). This helps the model distinguish between what is and isn’t covered by standard homeowners insurance.

```{r}
geometry <- st_geometry(model_ready)  
model_ready_no_geom <- model_ready %>%
  st_drop_geometry()

sev_weather_property_cols <- names(model_ready_no_geom) %>%
  str_subset("damage_property_(tornado|excessive_heat|heavy_snow|high_wind|hail|winter_storm|blizzard|ice_storm|strong_wind|lightning)")

sev_weather_crops_cols <- names(model_ready_no_geom) %>%
  str_subset("damage_crops_(tornado|excessive_heat|heavy_snow|high_wind|hail|winter_storm|blizzard|ice_storm|strong_wind|lightning)")

disaster_property_cols <- names(model_ready_no_geom) %>%
  str_subset("damage_property_(hurricane|flood|flash_flood|storm_surge_tide|wildfire|coastal_flood|lakeshore_flood|tsunami)")

disaster_crops_cols <- names(model_ready_no_geom) %>%
  str_subset("damage_crops_(hurricane|flood|flash_flood|storm_surge_tide|wildfire|coastal_flood|lakeshore_flood|tsunami)")

model_ready_no_geom <- model_ready_no_geom %>%
  mutate(
    num_sev_weather_property = rowSums(select(., all_of(sev_weather_property_cols)), na.rm = TRUE),
    num_sev_weather_crops = rowSums(select(., all_of(sev_weather_crops_cols)), na.rm = TRUE),
    disasters_property = rowSums(select(., all_of(disaster_property_cols)), na.rm = TRUE),
    disasters_crops = rowSums(select(., all_of(disaster_crops_cols)), na.rm = TRUE)
  ) %>%
  mutate(
    home_damage = num_sev_weather_property + num_sev_weather_crops,
    disaster_damage = disasters_property + disasters_crops,
    combined_damage = home_damage + disaster_damage + 2000,
    climate_index = combined_damage / RiskIndex,
    building_risk = (ExpAnnualLoss / Total_Value) * PriceDef,
    realized_damaged = (combined_damage / ExpAnnualLoss) * RiskIndex,
    actual_NRI = (combined_damage) * (Social/ Community)
  )

model_ready <- st_sf(model_ready_no_geom, geometry = geometry)
```

```{r}
states <- c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", 
            "Delaware", "Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", 
            "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", "Maryland", 
            "Massachusetts", "Michigan", "Minnesota", "Mississippi", "Missouri", "Montana", 
            "Nebraska", "Nevada", "New Hampshire", "New Jersey", "New Mexico", "New York", 
            "North Carolina", "North Dakota", "Ohio", "Oklahoma", "Oregon", "Pennsylvania", 
            "Rhode Island", "South Carolina", "South Dakota", "Tennessee", "Texas", 
            "Utah", "Vermont", "Virginia", "Washington", "West Virginia", "Wisconsin", 
            "Wyoming")

square_miles <- c(52420, 665384, 113990, 53179, 163695, 104094, 5543, 2489, 65758, 59425, 10932, 83569, 
                  57914, 36420, 56273, 82278, 40408, 52378, 35380, 12406, 10554, 96714, 86936, 48432, 
                  69707, 147040, 77348, 110572, 9349, 8723, 121590, 54555, 53819, 70698, 44826, 
                  69899, 98379, 46054, 1545, 32020, 77116, 42144, 268596, 84897, 9616, 42775, 71298, 
                  24230, 65496, 97813) # Census Bureau 

years <- rep(2008:2019, each = length(states))

state_area <- data.frame(State = rep(states, times = length(2008:2019)),
                       Square_Miles = rep(square_miles, times = length(2008:2019)),
                       Year = years)

state_area <- state_area[!state_area$State %in% c("Alaska", "Hawaii"), ]

model_ready <- model_ready %>% 
  left_join(state_area, by = c("State", "Year"))

model_ready$PopDensity <- model_ready$Total_Residents / model_ready$Square_Miles

model_ready <- model_ready %>%
  mutate(Region = case_when(
    State %in% c("Minnesota", "Iowa", "Missouri", "Wisconsin", "Indiana", "Illinois", "Michigan", "Ohio") ~ "Midwest",
    State %in% c("Washington", "Oregon", "Idaho") ~ "Northwest",
    State %in% c("Colorado", "Kansas", "Montana", "North Dakota", "Nebraska", "South Dakota", "Wyoming") ~ "North Central",
    State %in% c("Connecticut", "Delaware", "Kentucky", "Massachusetts", "Maryland", "Maine", "New Hampshire", "New Jersey", "New York", "Pennsylvania", "Rhode Island", "Virginia", "Vermont", "West Virginia") ~ "Northeast",
    State %in% c("Arizona", "Nevada", "Utah", "California") ~ "Southwest",
    State %in% c("Oklahoma", "Texas", "New Mexico", "Louisiana") ~ "South Central",
    State %in% c("North Carolina", "South Carolina", "Georgia", "Alabama", "Mississippi", "Florida", "Tennessee", "Arkansas") ~ "Southeast",
    TRUE ~ "Unknown"  # Default case for missing or unmatched states
  ))

model_ready <- model_ready %>%
  arrange(Year, State)

### STATE-YEAR FIXED EFFECTS ###
model_ready <- model_ready %>% 
  mutate(RegionYearFE = paste(Region,Year, sep= "-"))
```

```{r}
model_ready_scaled <- model_ready %>%
  mutate(across(c(homePrems, PriceDef, disaster_damage, PopDensity, td_per_cap, MedHomeIncome, home_damage, Percent_white, AvgHomePrice, ExpAnnualLoss, RiskIndex, Total_Buidling_Value, combined_damage, climate_index, building_risk, realized_damaged, actual_NRI), 
                ~ {
                  scaled_value <- (.-min(.)) / (max(.) - min(.))
                  scaled_value[scaled_value == 0] <- 0.00000000000001
                  scaled_value
                }))
```

# Neighors and Weights

Based on my intuition, a dynamic weight system seems most appropriate for capturing the effects of storms on homeowners' premiums. Initially, I used a queen contiguity approach with equal weighting for all neighbors, but that didn't seem suitable. A massive cold storm will likely affect the entire Northeast, but not evenly. So, I switched to distance-based neighbors, setting a 720 km threshold to ensure Texas and Louisiana were considered neighbors. Looking at states like New York and Alabama, the assigned neighbors seemed reasonable. The only downside is that California ends up with just one neighbor, but I think the benefits of properly fitting the Northeast outweigh that drawback.

```{r}
nb <- model_ready %>%
  filter(Year %in% c(2011))

centroids <- st_centroid(nb)  # This should already be in 'sf' format

centroids_sp <- as(centroids, "Spatial")

dist_matrix <- spDists(centroids_sp)

# Create neighbors list
neighbors <- dnearneigh(centroids_sp, 0, 550)

# Define row indices for California, Arizona, and Oregon
california_index <- 4L  # Ensure it's an integer
arizona_index <- 2L
oregon_index <- 35L

neighbors[[california_index]] <- sort(unique(as.integer(c(neighbors[[california_index]], arizona_index, oregon_index))))
neighbors[[arizona_index]] <- sort(unique(as.integer(c(neighbors[[arizona_index]], california_index))))
neighbors[[oregon_index]] <- sort(unique(as.integer(c(neighbors[[oregon_index]], california_index))))

# Define row indices for Florida and its neighboring states
florida_index <- 8L  # Replace X with Florida's row index
mississippi_index <- 22L  # Replace Y with Mississippi's row index
alabama_index <- 1L   # Replace Z with Alabama's row index
south_carolina_index <- 38L  # Replace A with South Carolina's row index
louisiana_index <- 16L  # Replace B with Louisiana's row index

# Update the neighbors list
neighbors[[florida_index]] <- sort(unique(as.integer(c(neighbors[[florida_index]], mississippi_index, alabama_index, south_carolina_index, louisiana_index))))
neighbors[[mississippi_index]] <- sort(unique(as.integer(c(neighbors[[mississippi_index]], florida_index))))
neighbors[[alabama_index]] <- sort(unique(as.integer(c(neighbors[[alabama_index]], florida_index))))
neighbors[[south_carolina_index]] <- sort(unique(as.integer(c(neighbors[[south_carolina_index]], florida_index))))
neighbors[[louisiana_index]] <- sort(unique(as.integer(c(neighbors[[louisiana_index]], florida_index))))

# Define row indices for Texas and its neighboring states
texas_index <- 41L  # Replace X with Texas's row index
new_mexico_index <- 29L  # Replace Y with New Mexico's row index
arkansas_index <- 3L  # Replace A with Arkansas's row index

# Update the neighbors list
neighbors[[texas_index]] <- sort(unique(as.integer(c(neighbors[[texas_index]], new_mexico_index, louisiana_index, arkansas_index))))
neighbors[[new_mexico_index]] <- sort(unique(as.integer(c(neighbors[[new_mexico_index]], texas_index))))
neighbors[[louisiana_index]] <- sort(unique(as.integer(c(neighbors[[louisiana_index]], texas_index))))
neighbors[[arkansas_index]] <- sort(unique(as.integer(c(neighbors[[arkansas_index]], texas_index))))

# Define row indices for Colorado, Kansas, and Nebraska
colorado_index <- 5L  # Replace X with Colorado's row index
kansas_index <- 14L  # Replace Y with Kansas's row index
nebraska_index <- 25L  # Replace Z with Nebraska's row index

# Update the neighbors list
neighbors[[colorado_index]] <- sort(unique(as.integer(c(neighbors[[colorado_index]], kansas_index, nebraska_index))))
neighbors[[kansas_index]] <- sort(unique(as.integer(c(neighbors[[kansas_index]], colorado_index))))
neighbors[[nebraska_index]] <- sort(unique(as.integer(c(neighbors[[nebraska_index]], colorado_index))))

# Define row indices for Idaho, Utah, and Wyoming
idaho_index <- 10L  # Replace X with Idaho's row index
utah_index <- 42L  # Replace Y with Utah's row index
wyoming_index <- 48L  # Replace Z with Wyoming's row index

# Update the neighbors list
neighbors[[idaho_index]] <- sort(unique(as.integer(c(neighbors[[idaho_index]], utah_index, wyoming_index))))
neighbors[[utah_index]] <- sort(unique(as.integer(c(neighbors[[utah_index]], idaho_index, wyoming_index))))
neighbors[[wyoming_index]] <- sort(unique(as.integer(c(neighbors[[wyoming_index]], idaho_index, utah_index))))


# Update the neighbors list
neighbors[[nebraska_index]] <- sort(unique(as.integer(c(neighbors[[nebraska_index]], kansas_index, wyoming_index))))
neighbors[[kansas_index]] <- sort(unique(as.integer(c(neighbors[[kansas_index]], nebraska_index))))
neighbors[[wyoming_index]] <- sort(unique(as.integer(c(neighbors[[wyoming_index]], nebraska_index))))


# Define row indices for Washington and Idaho
washington_index <- 45L  # Replace X with Washington's row index

# Update the neighbors list
neighbors[[washington_index]] <- sort(unique(as.integer(c(neighbors[[washington_index]], idaho_index))))
neighbors[[idaho_index]] <- sort(unique(as.integer(c(neighbors[[idaho_index]], washington_index))))

# Define row indices for Nevada and Idaho
nevada_index <- 26L  # Replace X with Nevada's row index

# Update the neighbors list
neighbors[[nevada_index]] <- sort(unique(as.integer(c(neighbors[[nevada_index]], idaho_index))))
neighbors[[idaho_index]] <- sort(unique(as.integer(c(neighbors[[idaho_index]], nevada_index))))

# Update the neighbors list
neighbors[[new_mexico_index]] <- sort(unique(as.integer(c(neighbors[[new_mexico_index]], arizona_index))))
neighbors[[arizona_index]] <- sort(unique(as.integer(c(neighbors[[arizona_index]], new_mexico_index))))

# Define row indices for Montana, South Dakota, and North Dakota
montana_index <- 24L  # Replace X with Montana's row index
south_dakota_index <- 39L  # Replace Y with South Dakota's row index
north_dakota_index <- 32L # Replace Z with North Dakota's row index

# Update the neighbors list
neighbors[[montana_index]] <- sort(unique(as.integer(c(neighbors[[montana_index]], south_dakota_index, north_dakota_index))))
neighbors[[south_dakota_index]] <- sort(unique(as.integer(c(neighbors[[south_dakota_index]], montana_index, north_dakota_index))))
neighbors[[north_dakota_index]] <- sort(unique(as.integer(c(neighbors[[north_dakota_index]], montana_index, south_dakota_index))))

# Update the neighbors list
neighbors[[wyoming_index]] <- sort(unique(as.integer(c(neighbors[[wyoming_index]], south_dakota_index))))
neighbors[[south_dakota_index]] <- sort(unique(as.integer(c(neighbors[[south_dakota_index]], wyoming_index))))

# Define row indices for Kentucky and Virginia
kentucky_index <- 15L  # Replace X with Kentucky's row index
virginia_index <- 44L  # Replace Y with Virginia's row index

# Update the neighbors list
neighbors[[kentucky_index]] <- sort(unique(as.integer(c(neighbors[[kentucky_index]], virginia_index))))
neighbors[[virginia_index]] <- sort(unique(as.integer(c(neighbors[[virginia_index]], kentucky_index))))

# Define row indices for Tennessee and North Carolina
tennessee_index <- 40  # Replace Z with Tennessee's row index
north_carolina_index <- 31L  # Replace A with North Carolina's row index

# Update the neighbors list
neighbors[[tennessee_index]] <- sort(unique(as.integer(c(neighbors[[tennessee_index]], north_carolina_index))))
neighbors[[north_carolina_index]] <- sort(unique(as.integer(c(neighbors[[north_carolina_index]], tennessee_index))))

# Update the neighbors list
neighbors[[utah_index]] <- sort(unique(as.integer(c(neighbors[[utah_index]], arizona_index))))
neighbors[[arizona_index]] <- sort(unique(as.integer(c(neighbors[[arizona_index]], utah_index))))

# Update the neighbors list
neighbors[[arkansas_index]] <- sort(unique(as.integer(c(neighbors[[arkansas_index]], tennessee_index))))
neighbors[[tennessee_index]] <- sort(unique(as.integer(c(neighbors[[tennessee_index]], arkansas_index))))

# Define row indices for Missouri and Kentucky
missouri_index <- 23L  # Replace Z with Missouri's row index

# Update the neighbors list
neighbors[[missouri_index]] <- sort(unique(as.integer(c(neighbors[[missouri_index]], kentucky_index))))
neighbors[[kentucky_index]] <- sort(unique(as.integer(c(neighbors[[kentucky_index]], missouri_index))))

# Define row indices for Michigan and Indiana
michigan_index <- 20L  # Replace X with Michigan's row index
indiana_index <- 12L  # Replace Y with Indiana's row index

# Update the neighbors list
neighbors[[michigan_index]] <- sort(unique(as.integer(c(neighbors[[michigan_index]], indiana_index))))
neighbors[[indiana_index]] <- sort(unique(as.integer(c(neighbors[[indiana_index]], michigan_index))))

# Define row indices for Michigan and Illinois
illinois_index <- 11L  # Replace Y with Illinois's row index

# Update the neighbors list
neighbors[[michigan_index]] <- sort(unique(as.integer(c(neighbors[[michigan_index]], illinois_index))))
neighbors[[illinois_index]] <- sort(unique(as.integer(c(neighbors[[illinois_index]], michigan_index))))

iowa_index <- 13L  # Replace Y with Iowa's row index

# Update the neighbors list
neighbors[[south_dakota_index]] <- sort(unique(as.integer(c(neighbors[[south_dakota_index]], iowa_index))))
neighbors[[iowa_index]] <- sort(unique(as.integer(c(neighbors[[iowa_index]], south_dakota_index))))

# Define row indices for Nevada, Arizona, and Oregon
oregon_index <- 35L  # Replace Z with Oregon's row index

# Update the neighbors list
neighbors[[arizona_index]] <- sort(unique(as.integer(c(neighbors[[arizona_index]], nevada_index))))  # Arizona only neighbors with Nevada
neighbors[[nevada_index]] <- sort(unique(as.integer(c(neighbors[[nevada_index]], arizona_index, oregon_index))))  # Nevada neighbors with Arizona & Oregon
neighbors[[oregon_index]] <- sort(unique(as.integer(c(neighbors[[oregon_index]], nevada_index))))  # Oregon neighbors with Nevada





lw <- nb2listw(neighbors, style = "B", zero.policy = TRUE)  # Use binary weights first

# Inverse distance weighting
for (i in 1:length(neighbors)) {
  neighbor_indices <- neighbors[[i]]
  total_weight <- 0
  for (j in 1:length(neighbor_indices)) {
    distance <- dist_matrix[i, neighbor_indices[j]]
    if (!is.na(distance) && distance != 0) {
      weight <- 1 / distance
      lw$weights[[i]][j] <- weight
      total_weight <- total_weight + weight
    } else {
      lw$weights[[i]][j] <- 0
    }
  }
  # Normalize weights
  if (total_weight != 0) {
    lw$weights[[i]] <- lw$weights[[i]] / total_weight
  }
}

# Function to replicate weights and neighbors while maintaining class attributes and style (NOT USED FOR SPATIAL PLM)
replicate_listw <- function(lw, times) {
  replicated_weights <- rep(lw$weights, times)
  replicated_neighbours <- do.call("c", replicate(times, lw$neighbours, simplify = FALSE))
  
  # Ensure neighbours has correct nb structure and region.id is a character
  class(replicated_neighbours) <- class(lw$neighbours)
  
  # create listw structure, copying the structure of lw
  expanded_weights <- list(
    style = lw$style,  # Place style first
    neighbours = replicated_neighbours,  # Place neighbours second
    weights = replicated_weights  # Place weights last
  )
  class(expanded_weights) <- class(lw)
  
  return(expanded_weights)
}

# Determine the number of years
num_years <- length(unique(final_df$Year))
num_rows <- nrow(final_df)
num_reps <- num_rows / length(lw$weights)

weights_across_years <- replicate_listw(lw, num_reps)

# Plot neighbors connections
plot(st_geometry(nb), border = "black") # plot polygons with gray borders
plot(neighbors, st_coordinates(st_centroid(st_geometry(nb))), add = TRUE, col = "blue", lwd = 1.5)
```

```{r}
lm2 <- lm(rentersPrems ~ log(home_damage) + log(disaster_damage) + AvgHomePrice + Percent_white  + log(PopDensity), data = model_ready_scaled)

summary(lm2)

### Moran's I ### which quantifies how similar each region is with its neighbors and averages all these assessments.
residuals <- residuals(lm2)

moran.test(residuals, weights_across_years, alternative = "greater")

MC <- moran.mc(residuals, weights_across_years, nsim = 999, alternative = "greater") # prefered method 

plot(MC)

LM_state <- lm.LMtests(lm2, weights_across_years, test = "all")
LM_state
```

# Spatial Models
```{r}
panel_data <- model_ready 

# Scale variables
panel_data_scaled <- panel_data %>%
  mutate(across(c(homePrems, rentersPrems, HealthcareExp, Pop18_64, Uninsured, Medicaid, PriceDef, disaster_damage, PopDensity, Total_Residents, td_per_cap, MedHomeIncome, home_damage, Percent_white, AvgHomePrice, ExpAnnualLoss, RiskIndex, Total_Buidling_Value, combined_damage, climate_index, building_risk, realized_damaged, actual_NRI), 
                ~ {
                  scaled_value <- (.-min(.)) / (max(.) - min(.))
                  scaled_value[scaled_value == 0] <- 0.00000000000001
                  scaled_value
                }))
```

```{r}
## Final Model
# Error
M1S <- spml(rentersPrems ~ log(home_damage) + lag(log(home_damage)) + log(disaster_damage) + lag(log(disaster_damage)) + AvgHomePrice + Percent_white + log(PopDensity),
                   data = panel_data_scaled, 
                   index = c("State", "Year"),
                   listw = lw,
                   model = "within",
                   effect = "individual",
                   lag = TRUE,
                   spatial.error = "none")

summary_M1S <- summary(M1S)

summary_M1S

model1 <- coeftest(summary_M1S)
```

```{r}
## Final Model
# Error
M1Y <- spml(rentersPrems ~ log(home_damage) + lag(log(home_damage)) + log(disaster_damage) + lag(log(disaster_damage)) + AvgHomePrice + Percent_white  + log(PopDensity),
                   data = panel_data_scaled, 
                   index = c("State", "Year"),
                   listw = lw,
                   model = "within",
                   effect = "time",
                   lag = TRUE,
                   spatial.error = "none")

summary_M1Y <- summary(M1Y)

summary_M1Y

model2 <- coeftest(summary_M1Y)
```


